---
title: "Preparing the Data"
output: html_notebook
---

# Packages

```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(xts)
library(PerformanceAnalytics)
library(lubridate)

`%ni%` <- Negate(`%in%`)
```

# Import Data

Index return (IBX) and Risk free return (CDI):

```{r}
ind <- read.csv("indice.csv")
ind$Data <- as.Date(ind$Data)

rf <- read.csv("rf.csv")
rf$Data <- as.Date(rf$Data)

ind_rf <- ind %>%
  mutate(rf$Risk_free) %>%
  set_names("Data", "ret_ibx", "ret_rf")

rm(ind, rf)
```

Funds' information:

```{r}
daily_data <- readRDS("dados_tratados_diarios.rds")
names(daily_data) <- c("capt_liq", "capt", "cota", "n_cotistas", "patrim_liq", "resg", "tx_adm")
```

```{r}
cadastral_info <- readRDS("dados_cadastrais.rds")

# We transform our columns so it can be used in a model
cadastral_info <- cadastral_info %>%
  dplyr::select(codigo, invest_qualificado, alavancado, data_inicio, prazo_conv_resg, prazo_pag_resg, aplic_min_inic, forma_condominio, fundo_cotas, fundo_exclusivo, data_inicio, data_fim) %>%
  mutate(invest_qualificado = ifelse(invest_qualificado == "N達o", 0, 1),
         alavancado = ifelse(alavancado == "N達o", 0, 1),
         prazo_conv_resg = as.numeric(str_sub(prazo_conv_resg, -3, -1)),
         prazo_pag_resg = as.numeric(str_sub(prazo_pag_resg, -3, -1)),
         prazo_receb_resg = prazo_conv_resg + prazo_pag_resg,
         aplic_min_inic = as.numeric(aplic_min_inic),
         forma_condominio = ifelse(forma_condominio == "Fechado", 0, 1),
         fundo_cotas = ifelse(fundo_cotas == "N達o", 0, 1),
         fundo_exclusivo = ifelse(fundo_exclusivo == "N達o", 0, 1))

# Soma funds are in one database but not in another one
diff_cadastral_daily <- setdiff(cadastral_info$codigo, colnames(daily_data[["cota"]]))

# We keep only the funds that are in both databases
cadastral_info <- cadastral_info %>% dplyr::filter(codigo %ni% diff_cadastral_daily)
```

# Funds selection

* The fund needs to be active for at least 12 months;
* During these 12 months, the fund needs to have data for at least 1 day when the market was open;
* The fund cannot be an exclusive one.

```{r}
# start_pre_date = when the estimation date will start
# end_pre_date = when the estimation date will end

get_funds_eleg <- function(start_pre_date, end_pre_date) {
  # First and second restrictions
  funds_eleg <- cadastral_info %>%
    dplyr::filter(data_inicio <= start_pre_date & (data_fim >= end_pre_date | is.na(data_fim))) %>% # First
    dplyr::filter(fundo_exclusivo == 0) %>% # Second
    dplyr::select(codigo)

  # Third restriction
  funds_eleg <- daily_data[["cota"]] %>%
    dplyr::filter(Data >= start_pre_date & Data <= end_pre_date) %>%
    dplyr::select(all_of(funds_eleg$codigo)) %>%
    dplyr::select_if(~ !any(is.na(.)))

  funds_eleg <- colnames(funds_eleg)

  return(funds_eleg)
}
```

# Feature Creation

## Returns Statistics

```{r}
# ret_port = portfolio return
# ind = index return
# rf = risk free return

get_stat_ret <- function(ret_port, ind, rf) {

  # Annualized Cumulative Return
  ret_period <- function(ret) {
    prod(ret + 1) ^ (252 / length(ret)) - 1
  }

  # CAPM
  regres <- lm(I(ret_port - rf) ~ I(ind - rf))
  # Get regression coefficients
  regres_coef <- summary(regres)$coefficients

  # Jansen's Alpha
  alpha <- (regres_coef[1, 1] + 1) ^ 252 - 1
  # CAPM Beta
  beta <- regres_coef[2, 1]

  # Portfolio annualized cumulative return
  cumulative_ret <- ret_period(ret_port)

  # Annualized Portfolio volatility
  vol <- sd(ret_port) * sqrt(252)

  # Israelsen Sharpe Ratio and Information Ratio (ttps://doi.org/10.1057/palgrave.jam.2240158)
  ER <- mean(ret_port - rf)
  SD <- sd(ret_port - rf)
  MSR <- ER / (SD^(ER / abs(ER))) # Modified SR

  ER <- mean(ret_port - ind)
  SD <- sd(ret_port - ind)
  MIR <- ER / (SD^(ER / abs(ER))) # Modified IR

  # CVaR
  cvar <- CVaR(ret_port, method = "historical")

  results <- data.frame(
    cumulative_ret, vol, MSR, MIR,
    beta, alpha, cvar
  )

  return(results)
}
```

```{r}
# ret_port = portfolio return
# ind = index return
# rf = risk free return

get_stat_ret <- function(ret_port, ind, rf) {

  # Annualized Cumulative Return
  ret_period <- function(ret) {
    prod(ret + 1) ^ (252 / length(ret)) - 1
  }

  # CAPM
  regres <- lm(I(ret_port - rf) ~ I(ind - rf))
  # Get regression coefficients
  regres_coef <- summary(regres)$coefficients

  # Jansen's Alpha
  alpha <- (regres_coef[1, 1] + 1) ^ 252 - 1
  # CAPM Beta
  beta <- regres_coef[2, 1]

  # Portfolio annualized cumulative return
  cumulative_ret <- ret_period(ret_port)
  
  # Index annualized cumulative return
  cumulative_ind <- ret_period(ind)

  # Annualized Portfolio volatility
  vol <- sd(ret_port) * sqrt(252)
  
  # Annualized Index volatility
  vol_ind <- sd(ret_port) * sqrt(252)
  
  # Tracking Error
  track_error <- sd(ret_port - ind) * sqrt(252)

  # Israelsen Sharpe Ratio and Information Ratio (ttps://doi.org/10.1057/palgrave.jam.2240158)
  ER <- mean(ret_port - rf)
  SD <- sd(ret_port - rf)
  MSR <- ER / (SD^(ER / abs(ER))) # Modified SR

  ER <- mean(ret_port - ind)
  SD <- sd(ret_port - ind)
  MIR <- ER / (SD^(ER / abs(ER))) # Modified IR

  # CVaR
  cvar <- CVaR(ret_port, method = "historical")

  results <- data.frame(
    cumulative_ret, vol, 
    track_error, MIR, MSR,
    beta, alpha, cvar,
    cumulative_ind, vol_ind
  )

  return(results)
}
```


## Fund's Statistics

```{r}
# start_pre_date = when the estimation date will start
# end_pre_date = when the estimation date will end
# funds_eleg = which funds meet all the restrictions (result from "get_funds_eleg")

get_funds_stat <- function(start_pre_date, end_pre_date, funds_eleg) {
  # Return Statistics
  funds_returns <- daily_data[["cota"]] %>%
    dplyr::filter(Data >= start_pre_date & Data <= end_pre_date) %>%
    dplyr::select(all_of(funds_eleg))

  ind_rf_pre <- ind_rf %>% dplyr::filter(Data >= start_pre_date & Data <= end_pre_date)

  funds_stats <- lapply(funds_returns, function(x) get_stat_ret(x, ind_rf_pre$ret_ibx, ind_rf_pre$ret_rf))

  funds_stats <- do.call("rbind", funds_stats)

  # Assets Under Management
  funds_aum <- daily_data[["patrim_liq"]] %>%
    dplyr::filter(Data >= start_pre_date & Data <= end_pre_date) %>%
    dplyr::select(all_of(funds_eleg))

  funds_aum <- lapply(funds_aum, function(x) mean(x, na.rm = TRUE))

  funds_aum <- as.data.frame(do.call("rbind", funds_aum)) %>% set_names("aum")

  # Inflow
  funds_inflow <- daily_data[["capt"]] %>%
    dplyr::filter(Data >= start_pre_date & Data <= end_pre_date) %>%
    dplyr::select(all_of(funds_eleg))

  funds_inflow <- lapply(funds_inflow, function(x) sum(x, na.rm = TRUE))

  funds_inflow <- as.data.frame(do.call("rbind", funds_inflow)) %>% set_names("inflow")

  # Outflow
  funds_outflow <- daily_data[["resg"]] %>%
    dplyr::filter(Data >= start_pre_date & Data <= end_pre_date) %>%
    dplyr::select(all_of(funds_eleg))

  funds_outflow <- lapply(funds_outflow, function(x) sum(x, na.rm = TRUE))

  funds_outflow <- as.data.frame(do.call("rbind", funds_outflow)) %>% set_names("outflow")

  # Number of shareholders
  funds_share_hold <- daily_data[["n_cotistas"]] %>%
    dplyr::filter(Data >= start_pre_date & Data <= end_pre_date) %>%
    dplyr::select(all_of(funds_eleg))

  funds_share_hold <- lapply(funds_share_hold, function(x) mean(x, na.rm = TRUE))

  funds_share_hold <- as.data.frame(do.call("rbind", funds_share_hold)) %>% set_names("shareholders")

  # Join all data frames
  merge.all <- function(x, ..., by = "row.names") {
    L <- list(...)
    for (i in seq_along(L)) {
      x <- merge(x, L[[i]], by = by)
      rownames(x) <- x$Row.names
      x$Row.names <- NULL
    }
    return(x)
  }

  funds_stats <- merge.all(funds_stats, funds_aum, funds_inflow, funds_outflow, funds_share_hold)

  return(funds_stats)
}
```

# Dependent Variable (0 or 1)

```{r}
# start_eval_date = when the evaluation date will start
# end_eval_date = when the evaluation date will end
# funds_eleg = which funds meet all the restrictions (result from "get_funds_eleg")

get_dependent_var <- function(start_eval_date, end_eval_date, funds_eleg, type) {
  if(type == 'classif_closed'){
      # Determine which funds closed between the start and end of the evaluation period
    funds_closed <- cadastral_info %>%
      dplyr::filter(codigo %in% funds_eleg) %>%
      mutate(closed = ifelse(data_fim < start_eval_date | data_fim > end_eval_date | is.na(data_fim), 0, 1)) %>%
      dplyr::select(codigo, closed) %>%
      `rownames<-`(.$codigo) %>%
      dplyr::select(!codigo) %>% 
      set_names('dep_var') # The name must be the same for generalization purposes
    
    return(funds_closed)

  } else if(type == 'classif_outperfomance') {
     funds_returns_eval <- daily_data[['cota']] %>% dplyr::filter(Data > start_eval_date & Data <= end_eval_date) %>%
      dplyr::select(all_of(funds_eleg)) 
    
    ind_rf_eval <- ind_rf %>% dplyr::filter(Data > start_eval_date & Data <= end_eval_date)
    
    funds_stats_eval <- lapply(funds_returns_eval, function(x) prod(1 + (x - ind_rf_eval$ret_ibx), na.rm = TRUE) - 1)
    
    funds_stats_eval <- do.call('c', funds_stats_eval)
    
    funds_stats_eval <- as.data.frame(ifelse(funds_stats_eval > 0, 1, 0)) %>% set_names('dep_var') # The name must be the same for generalization purposes
    
    return(funds_stats_eval)
  } else if(type == 'regres_flows'){
    funds_flow <- daily_data[['capt_liq']] %>% dplyr::filter(Data >= start_eval_date & Data <= end_eval_date) %>%
        dplyr::select(all_of(funds_eleg)) 
    
    funds_flow <- lapply(funds_flow, function(x) sum(x, na.rm = TRUE))
    
    funds_flow <- as.data.frame(do.call('rbind', funds_flow)) %>% set_names('dep_var') # The name must be the same for generalization purposes 
    
    return(funds_flow)
  } else if(type == 'classif_flows'){
    funds_flow <- daily_data[['capt_liq']] %>% dplyr::filter(Data >= start_eval_date & Data <= end_eval_date) %>%
      dplyr::select(all_of(funds_eleg)) 
    
    funds_flow <- lapply(funds_flow, function(x) sum(x, na.rm = TRUE))
    
    funds_flow <- as.data.frame(do.call('rbind', funds_flow)) %>% 
      set_names('dep_var') %>% # The name must be the same for generalization purposes
      mutate(dep_var = ifelse(dep_var > 0, 1, 0))
    
    return(funds_flow)
  } else if(type == 'all'){
    types <- list('classif_closed', 'classif_outperfomance', 'regres_flows', 'classif_flows')
    
    funds_info <- lapply(types, function(x) get_dependent_var(start_eval_date, end_eval_date, funds_eleg, x))
    
    funds_info <- do.call('cbind', funds_info) %>%
      set_names('closed', 'outperf', 'regres_flows', 'classif_flows')

    return(funds_info)
  }

}
```

# Wrapper Function

```{r}
get_final_df <- function(start_pre_date, end_pre_date, start_eval_date, end_eval_date, type) {
  
  # Which funds meet all the requirements?
  funds_eleg <- get_funds_eleg(start_pre_date, end_pre_date)

  # Get the data regarding each funds
  funds_ret_stat <- get_funds_stat(start_pre_date, end_pre_date, funds_eleg)

  # Calculate how old was the fund at the end of the estimation period and get the cadastral information
  funds_cadastral <- cadastral_info %>%
    dplyr::filter(codigo %in% funds_eleg) %>%
    mutate(age = time_length(difftime(as.Date(end_pre_date), as.Date(data_inicio)), "years")) %>% # How old
    `rownames<-`(.$codigo) %>%
    dplyr::select(!c(codigo, data_inicio, data_fim, prazo_conv_resg, prazo_pag_resg)) # cadastral information

  # Get our dependent variable
  funds_dep_variable <- get_dependent_var(start_eval_date, end_eval_date, funds_eleg, type)

  funds_stat <- merge(funds_dep_variable, funds_ret_stat, by = 0) %>%
    `rownames<-`(.$Row.names) %>%
    dplyr::select(!Row.names)

  funds_stat <- merge(funds_stat, funds_cadastral, by = 0) %>%
    `rownames<-`(.$Row.names) %>%
    dplyr::select(!Row.names)

  return(funds_stat)
}
```

# Get clean data 

```{r, message=FALSE, warning=FALSE}
# How many months should we look back to calculate the statistics?
look_back <- 12
# How many months should we consider for making a prediction?
pred_n_months <- 12
# With what frequency (in months) should we make a new prediction?
rebalance_n_months <- 1
# When to start making a prediction
start_eval_date <- as.Date("2017-01-01")
# What type of dependent variable we want? ('all' with we want all of them)
type = 'all'

n_months <- interval(start_eval_date %m-% days(1), as.Date("2021-12-31")) %/% months(rebalance_n_months)
final_df <- vector("list", length = n_months)
for (i in 1:n_months) {
  start_eval_date <- as.Date(start_eval_date) %m+% months(rebalance_n_months)
  end_estim_date <- start_eval_date %m-% days(1)
  start_estim_date <- as.Date(end_estim_date) %m-% months(look_back)
  end_eval_date <- as.Date(start_eval_date) %m+% months(pred_n_months) %m-% days(1)

  final_df[[i]] <- get_final_df(start_estim_date, end_estim_date, start_eval_date, end_eval_date, type) %>%
    dplyr::select(!c(invest_qualificado, aplic_min_inic, fundo_exclusivo)) %>%
    mutate_at(vars(forma_condominio, fundo_cotas), ~ ifelse(is.na(.x), 0, .x)) %>%
    mutate(year = year(end_estim_date), month = month(end_estim_date))
}

final_df <- do.call("rbind", final_df)
```

```{r}
write.csv(final_df, 'data_modelling.csv')
```

